Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-15 main.tex
typo
CP progress on basic definition

\bcp{Idea: Remove the HIGH/LOW part from contours and instead make it part
  of the $X$ parameter to both properties and policies.  (I.e., we will have
  ``X properties'' and ``X policies'' for each X.)}

A {\em contour}
$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}\times
\{\OBS,\HIDDEN\}$ is a map from machine components to pairs of ``security
levels,'' where $\HIGH$ marks locations that may contain secret or
low-integrity data and $\LOW$ marks those that must only ever contain
publicly readable or high-integrity data, and ``observabilities,'' where
$\OBS$ marks locations whose contents are observable from the outside (by
the ``attacker,'' e.g., by DMA) and $\HIDDEN$ marks internal ``hidden''
state that cannot be observed directly (e.g., processor registers and memory
addresses for temporary variables will usually be hidden).

A {\em contoured trace} is a finite or infinite sequence
$\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs \bcp{such
  that the machine parts agree with the full step}.

@@

\bcp{Try it the other way round: a sequence of maps from components to
  values.  Need to extend values with $\bot$ for always-hidden elements.
  Emphasize that ``observability is not observable.''}
\bcp{Also, change OBS to VIS}

A {\em component observation} is a finite or infinite sequence of words in
which adjacent elements are always distinct (``no stuttering'').  An {\em
  observation} is a map from machine components to component observations.

The observation arising from a contoured trace $\seq{(\contour,\mach)}$,
written $\obstrace(\seq{(\contour,\mach)})$, is a map taking each machine
component to the (finite or infinite) sequence of the {\em distinct} values
taken on by this component at moments when it is {\em observable}.

Formally, $\obstrace(\seq{(\contour,\mach)})$ can be constructed as
follows. For each component $\component$,
%
if
$\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
$\obstrace(\component) = [\,]$;
%
otherwise, $\obstrace(\component)$ is a sequence whose first element
is $\mach_i(\component)$ for the smallest $i$ such that
$\contour_i(\component) = (\underscore,\OBS)$ and whose $(m+1)^{th}$ element
(if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
than the index $i$ of the $m^{th}$ element and such that
$\contour_j(\component) = (\underscore,\OBS)$ and
$\mach_j(\component) \neq \mach_i(\component)$.
%
\bcp{Can be defined more simply}
\iftext
\bcp{Unfinished and horrible...  But maybe the textual description is OK?}
\newcommand{\aux}{\mathit{aux}}
\bcp{Don't like the ``rest'' argument and the fact that we don't
  have a metavariable for sequences of (contour, machine state) pairs.
  Consider defining one.}
%
\[
\begin{array}{lcl}
\aux \; ( [\,] ) \; \trace& = & \trace
\\
\aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
  & = &
        \aux \;( \mathit{rest} )
        \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
\\
\\
\obstrace \;( (\contour_0,\mach_0),\, \ldots )
  &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
      \component.\, \bot)
\end{array}
\]
\fi

@@

@@

@@

Here we consider stack protection as an instance of generalized interference \apt{or whatever it's called}.
Our notion of stack protection includes aspects both of control integrity and data security.
In the terminology of Skorstengaard et al.~\cite{Skorstengaard+19} we enforce:

>>>>

A {\em contour}
$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\VIS,\HID\}$ is a map
from machine components to ``observabilities,'' where
$\VIS$ marks locations whose contents are observable from the outside (by
the ``attacker,'' e.g., by DMA) and $\HID$ marks internal ``hidden''
state that cannot be observed directly (e.g., processor registers and memory
addresses for temporary variables will usually be hidden).

A {\em contoured trace} is a finite or infinite sequence
$\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs in which
adjacent machine states are related by $\stepstounderfull$---i.e., in which
$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the sequence.

@@

A {\em component observation} is either a word or the special value $\bot$.
A {\em state observation} is a map from machine components to component
observations.

An {\em observation} is a nonempty, finite or infinite sequence of state
observations in which (1) adjacent elements are always distinct (``no
stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
the the $j^{th}$ state observation in the sequence must also be mapped to
$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
component's value changes from $\bot$ to anything else, it never changes
back to $\bot$ again).

In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
to its value in $\mach_0$ and every hidden
component to $\bot$; 
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
visible component (where $\contour_{i+1} = \VIS$), copying its
value from $\mach_{i+1}$, and (2) for each
hidden component, where $\contour_i = \VIS$, copying its
value from $\stateobs_{i+1}$, and (3) if the resulting $\stateobs_{i+1}$ is
identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
\iftext\bcp{Too informal!!}\fi

% For each component $\component$,
% %
% if
% $\contour_i(\component) = (\underscore,\HID)$ for all $i$, then
% $\obstrace(\component) = [\,]$;
% %
% otherwise, $\obstrace(\component)$ is a sequence whose first element
% is $\mach_i(\component)$ for the smallest $i$ such that
% $\contour_i(\component) = (\underscore,\VIS)$ and whose $(m+1)^{th}$ element
% (if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
% than the index $i$ of the $m^{th}$ element and such that
% $\contour_j(\component) = (\underscore,\VIS)$ and
% $\mach_j(\component) \neq \mach_i(\component)$.
% %
% \bcp{Can be defined more simply}
% \iftext
% \bcp{Unfinished and horrible...  But maybe the textual description is OK?}
% \newcommand{\aux}{\mathit{aux}}
% \bcp{Don't like the ``rest'' argument and the fact that we don't
%   have a metavariable for sequences of (contour, machine state) pairs.
%   Consider defining one.}
% %
% \[
% \begin{array}{lcl}
% \aux \; ( [\,] ) \; \trace& = & \trace
% \\
% \aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
%   & = &
%         \aux \;( \mathit{rest} )
%         \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
% \\
% \\
% \obstrace \;( (\contour_0,\mach_0),\, \ldots )
%   &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
%       \component.\, \bot)
% \end{array}
% \]
% \fi

@@

\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

@@

\bcp{Needs a parameter now.}

@@

Here we consider stack protection as an instance of generalized noninterference \apt{or whatever it's called}.
Our notion of stack protection includes aspects both of control integrity and data security.
In the terminology of Skorstengaard et al.~\cite{Skorstengaard+19} we enforce:
